---
title: "Statistical inference – ToothGrowth data"
author: "Marija Zoldin"
date: "05.07.2015"
output: pdf_document
---

### Overview

The goal of this task was to load ToothGrowth data and apply learned knowledge from statistical inference class on Coursera. Tooth growth among six independent group will be examined.

### Exploratory analysis and summary statistics

Experiment was made on 60 guinea pigs. They were divided into 6 groups and each group was threated with different supplement dose(0.5,1 and 2mg) and delivery methods(OJ, AC).
Data set contains 60 observations (10 guinea pigs per each group) and three variable(Len, supp and dose). It was loaded via R with following code:

```{r}
data(ToothGrowth)
str(ToothGrowth)
```

Data set contains informations about 6 different groups per supplement dose and delivery methods.
Six independent groups needs to be compared by column len which represents Toothgrowth length.

New variable group was created – it represents each group of 10 pigs with different treatment:

```{r}
ToothGrowth$group<-"OJ_0.5";
ToothGrowth[ToothGrowth$supp=="OJ" & ToothGrowth$dose==1,]$group<-"OJ_1";
ToothGrowth[ToothGrowth$supp=="OJ" & ToothGrowth$dose==2,]$group<-"OJ_2";
ToothGrowth[ToothGrowth$supp=="VC" & ToothGrowth$dose==0.5,]$group<-"VC_0.5";
ToothGrowth[ToothGrowth$supp=="VC" & ToothGrowth$dose==1,]$group<-"VC_1";
ToothGrowth[ToothGrowth$supp=="VC" & ToothGrowth$dose==2,]$group<-"VC_2";
```

Mean value of the each group was plotted:

```{r}
library(ggplot2)  
t1<-aggregate(len~group, FUN=mean,data=ToothGrowth)
ggplot(t1,aes(group,len,fill=group))+geom_bar(stat="identity",position="dodge")+
  theme(axis.text.x =element_text(size  = 10,angle = 90,hjust = 1,vjust = 1))+
  scale_fill_discrete(name="Group",breaks=c("OJ_0.5", "OJ_1", "OJ_2","VC_0.5","VC_1","VC_2"),
  labels=c("Orange_0.5", "Orange_1", "Orange_2","VitaminC_0.5","VitaminC_1","VitaminC_2"))+ggtitle("Average mean of tooth length by group")
```

Summary statistics of length variable per each group was calculated:

```{r}
tapply(ToothGrowth$len, ToothGrowth$group, summary)
```

### Assumptions
How the groups are independent (unpaired – each 10 pigs have different treatment) and only 10 pigs per group are present t-test was applied (mostly applied when number of sample data set is very small). Unequal variance parameter was used because variances differ in groups. Variance calculation is presented below:

```{r}
tapply(ToothGrowth$len, ToothGrowth$group, var)
```

### Confidence intervals and t-test

Confidence intervals per groups(all possible groups were examined)  were calculated based on t.test formula. Here is an example for CI calculation for group 1(supp orange juice with 0.5 mg dose) and group 2 (supp orange juice with 1 mg dose):

```{r}
#take the subset
group1_2<-ToothGrowth[ToothGrowth$group %in% c("OJ_0.5","OJ_1"),]
#calculate CI
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group1_2)$conf
```

Below are calculations for all groups:

```{r}
group1_2<-ToothGrowth[ToothGrowth$group %in% c("OJ_0.5","OJ_1"),]
    #calculate CI
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group1_2)$conf
group1_3<-ToothGrowth[ToothGrowth$group %in% c("OJ_0.5","OJ_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group1_3)$conf
group1_4<-ToothGrowth[ToothGrowth$group %in% c("OJ_0.5","VC_0.5"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group1_4)$conf
group1_5<-ToothGrowth[ToothGrowth$group %in% c("OJ_0.5","VC_1"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group1_5)$conf
group1_6<-ToothGrowth[ToothGrowth$group %in% c("OJ_0.5","VC_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group1_6)$conf

group2_3<-ToothGrowth[ToothGrowth$group %in% c("OJ_1","OJ_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group2_3)$conf
group2_4<-ToothGrowth[ToothGrowth$group %in% c("OJ_1","VC_0.5"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group2_4)$conf
group2_5<-ToothGrowth[ToothGrowth$group %in% c("OJ_1","VC_1"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group2_5)$conf
group2_6<-ToothGrowth[ToothGrowth$group %in% c("OJ_1","VC_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group2_6)$conf

group3_4<-ToothGrowth[ToothGrowth$group %in% c("OJ_2","VC_0.5"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group3_4)$conf
group3_5<-ToothGrowth[ToothGrowth$group %in% c("OJ_2","VC_1"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group3_5)$conf
group3_6<-ToothGrowth[ToothGrowth$group %in% c("OJ_2","VC_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group3_6)$conf

group4_5<-ToothGrowth[ToothGrowth$group %in% c("VC_0.5","VC_1"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group4_5)$conf
group4_6<-ToothGrowth[ToothGrowth$group %in% c("VC_0.5","VC_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group4_6)$conf

group5_6<-ToothGrowth[ToothGrowth$group %in% c("VC_1","VC_2"),]
t.test(len ~ group, paired = FALSE, var.equal = FALSE, data = group5_6)$conf
```


### Conclusions

From the results above we can conclude that there is a statistically significant mean difference among all treatments – exceptions are groups  OJ(dose 1)  and Vitamin C (dose 2) and OJ and Vitamin C (both with dose 2). In confidence intervals that are calculated for those groups 0 is included in interval so we can assume that there is no difference in average tooth growth. Negative confidence intervals means that group 1 having slower tooth growth than the group2. Positive confidence interval tells us that group one have higher average tooth growth and if 0 is not included means that this positive value is statistically significant.

